### 2013-01-07 ISMB/introduction_ISMB.tex
Method... interrupted

We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}. Instead a local search approach, we propose to apply a global sampling strategy of the mutational landscape using our \RNAmutants algorithm \cite{Waldispuhl2008}. This methodology offers promising performance but  suffers from a prohibitive running time complexity. Following our work, Garcia-Martin and co-workers proposed with \RNAiFOLD an alternate methodology using constraint programming techniques to prune the mutational landscape. The latter also suffers from prohibitive running times, but it is worth noting that it proposes a seed-less approach to the RNA secondary structure design problem.

>>>>

We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}. Instead of a local search approach, we propose to apply a global sampling strategy of the mutational landscape using our \RNAmutants algorithm \cite{Waldispuhl2008}. This methodology offers promising performance but  suffers from a prohibitive running time complexity. Following our work, Garcia-Martin and co-workers proposed with \RNAiFOLD an alternate methodology using constraint programming techniques to prune the mutational landscape. The latter also suffers from prohibitive running times, but it is worth noting that it proposes a seed-less approach to the RNA secondary structure design problem.

### 2013-01-07 ISMB/methods_ISMB.tex
s
Method... interrupted

\newcommand{\PE}[1]{E(#1)}
\newcommand{\EI}{\text{EI}}
\newcommand{\ES}{\text{ES}}
\newcommand{\ISO}{\text{ISO}}
\newcommand{\Prob}{\text{P}}

We introduce a probabilistic model, aiming to design RNA sequences with a 
specific $\Gb+\Cb$ content, folding into a predefined 2D conformation.
We use a Boltzmann weighted distribution based on a pseudo-energy function 
$\ES(\cdot)$ taking into accounts stacked canonical base pairs free-energy.

%To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment. In this model, the probability that the nucleotide at a given position needs to be mutated (i.e. corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
%
%\subsection{Probabilistic model}
%Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
%then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
%\begin{align*}
%  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
%\end{align*}
%where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
%are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
%
%\subsubsection{Energy contribution}
\subsection{Definitions}
\subsubsection{Profile}
At every position, the nucleotide distribution is the same. 
It is first initialized equiprobably. A bisection scheme is then applied until the sampled sequences $\Cb+\Gb$ content is as desired. The probability
of a given nucleotide $a\in\B$ will be given by $\Prob(a)$ at every 
position.

\subsubsection{Energetic model}

The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}. Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by

\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}

where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e. not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$). This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).

%The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
%(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
%\subsubsection{Isostericity contribution}
%The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the contribution of the isostericity to the pseudo-energy as
%\begin{align*}
%  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
%	\frac{
%		\sum_{s'\in\Omega}
%			\text{\ISO}((s_i',s_j'),(a,b))}
%%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
%{		
%		|\Omega|
%	}
%\end{align*}
%is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
%The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Isostericity scores range between $0$ and $9.7$, with $0$ being assigned to a perfect isostericity, and a penalty of $10$ is used for missing entries.
%The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
%\subsection{Mutational profile of sequences}
%Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
%We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
%%  under the SCFG derivation $S$ 
%(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
%We define a variant of the
% \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
% desired probability,  the two  functions
%$\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$. 
%The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec}, is analogous to the \emph{inside} algorithm.
%It is the partition function, i.e. the sum of Boltzmann factors, over all sequences within $[i,j]$, 
%knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
%$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
% computes the \emph{outside} algorithm, i.e.  
%the partition function over sequences within $m$ mutations of $s$, restricted to two intervals $[0,i]\cup[j,n-1]$, and 
%knowing  that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$). A suitable combination of these terms, given in Equation~\eqref{eq:combine}, gives the total weight, and in turn the probability, of seeing a specific base at a given position.
%%Drawing a parallel with stochastic context-free grammars (SCFG), for which the inside-outside algorithm was introduced, the set of sequences can be seen as the language of words having length $n$, generated from a stochastic context-free grammar.
%%These rules would limit the number of mutations (i.e. )
%% $S$ can be considered as constraining weighing on the shape of parse trees the derivations in a classic 
%% of the {SCFG} generating  all  secondary structures of length  $|s|$. 
%Let $B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
%Given $s\in B^n$ an RNA sequence, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
%length $n$, and $S$ a secondary structure without pseudoknots. 
%Formally, if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities
% $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty 
% ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
% $[i,j]\subset[k,l])$. 
%Let us then remind a Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
%Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that

%\begin{equation}
%  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
%\end{equation}
%
%
%\subsubsection{Inside}
%\begin{figure}
%\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
%\caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
%can be decomposed as a sequence over $[i+1,j]$ preceded by a, possibly mutated, base at $i$ 
%(Unpaired case), a sequence over $[i+1,j-1]$ surrounded by some base-pair (Stacking-pair case), 
%or as two sequences over $[i+1,k-1]$ and $[k+1,j]$, completed by some base-pair (General base-pairing case). 
%In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
%and locally instanciated bases.}
%\end{figure}
%
%The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e. the sum of Boltzmann factors, over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively). 
%Such terms can be defined by recurrence, for which the following initial conditions holds:
%
%\begin{equation}
%	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
%	\begin{array}{ll}
%		1 &\text{If } m = 0\\
%		0 &\text{Otherwise.}
%	\end{array}\right.
%\label{eq:Z_in}
%\end{equation}
%In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$. Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.

The main recursion is composed of four terms:
\begin{equation}
	\Z{i,j}{a,b}:=\left\{
  \begin{array}{ll}
  		\displaystyle
      \sum_{a'\in \B}  
      \Prob(a')
      \cdot\Z{i+1,j}{a',b} & \text{If }S_{i}=-1\\
      \displaystyle
      \sum_{a',b'\in \B^2}
			 \Prob(a')
			 \cdot\Prob(b')
			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 \cdot \Z{i+1,j-1}{a',b'}&
			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
			 \displaystyle
      \sum_{a',b'\in \B^2}
      \Prob(a')
			\cdot\Prob(b')
			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
      \cdot\Z{i+1,k-1}{a',b'}
      \cdot\Z{k+1,j}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
      0 &\text{Otherwise}
	\end{array}\right.
\label{eq:Z_rec}
\end{equation}
The cases can be broken down as follows:
\begin{description}
\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
any sequence consists in a nucleotide $a'$ at position  $i$, 
followed by a sequence over $[i+1,j]$ and having flanking nucleotides $a'$ and $b$.
\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$.
Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
then a stacking energy contribution is added. 
\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
only term contributing directly to the energy is the nucleotide distribution $\Prob(\cdot)$. 
Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
\end{description}

\subsection{Stochastic backtrack}

\begin{algorithm}
	\SetAlgoLined
	\lIf{$i>j$} \Return{$\varepsilon$};\tcp*[f]{Empty sequence}\\
	rand $\leftarrow$ Random$\left(\Z{i,j}{a,b}\right)$\;
	k $\leftarrow S_i$\;
	\If(\tcp*[f]{Unpaired}){$k = -1$}{
		\For{$a'\in\B$}{
			rand $\leftarrow$ rand $- \Prob(a')\cdot \Z{i+1,j}{a',b}$\;
			\lIf{$\text{rand}<0$}\Return{$a'\cdot\text{Backtrack}\left((i+1,j),(a',b),S\right)$}\;
		}
	}
	\Else{
		\If(\tcp*[f]{Stack}){$k=j \land S_{i-1}=j+1$}{
			\For{$(a',b')\in\B\times\B$}{
				rand $\leftarrow$ rand $-
			 	\Prob(a')
			 	\cdot\Prob(b')
			 	\cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
				\If{$\text{rand}<0$}{
					\Return{$a'\cdot\text{Backtrack}\left((i+1,j-1),(a',b'),S\right)\cdot b'$}\;		
				}
			}
		}
		\Else(\tcp*[f]{Paired not stacked}){
			\For{$(a',b')\in\B\times\B$}{
				rand $\leftarrow$ rand $-
				\Prob(a')
				\cdot\Prob(b')
				\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
  	    \cdot\Z{i+1,k-1}{a',b'}
    	  \cdot\Z{k+1,j}{b',b}$\;	
 				\If{$\text{rand}<0$}{
					\Return{$a'
					\cdot\text{Backtrack}\left((i+1,k-1),(a',b'),S\right)
					\cdot b'
					\cdot\text{Backtrack}\left((k+1,j),(b',b),S\right)
					$}\;	
				}
			}
		}
	}
\caption{Backtrack$\left((i,j),(a,b),S\right)$\label{alg:back}}
\end{algorithm}

>>>>

We introduce a probabilistic model for the design of RNA sequences with a specific $\Gb+\Cb$ content and folding into a predefined secondary structure.
For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
stacked canonical base-pairs.

We show how a modification of the dynamic programming scheme introduced in the previously introduced \RNAmutants allow for the sampling of good and diverse design candidates in linear time and space complexities.

%To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment. In this model, the probability that the nucleotide at a given position needs to be mutated (i.e. corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
%
%\subsection{Probabilistic model}
%Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
%then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
%\begin{align*}
%  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
%\end{align*}
%where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
%are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
%
%\subsubsection{Energy contribution}
\subsection{Definitions}

%\subsubsection{Profile}
%At every position, the nucleotide distribution is the same. 
%It is first initialized equiprobably. A bisection scheme is then applied until the sampled sequences $\Cb+\Gb$ content is as desired. The probability
%of a given nucleotide $a\in\B$ will be given by $\Prob(a)$ at every 
%position.

\subsubsection{Secondary structure}
Todo...

\subsubsection{Energy model}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}). Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e. not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$). This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = \infty$).

\subsection{\GCContent weighted pseudo-Boltzmann ensemble and distribution}

In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, let us introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences. Let $\Target$ be a targeted secondary structure and $\gc(s)$ be the number of occurrences of \Gb and \Cb in an RNA sequence $s$, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
\begin{equation}
\B_{x, \Target}(s) = e^{\frac{-\ES(s,\Target)}{RT}}\cdot x^{\gc(s)}
\label{def:genBoltz}
\end{equation}
where $R$ is the Boltzmann constant and $T$ the temperature in Kelvin.

Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$ gives the pseudo-partition function $\mathcal{Z}_{\Target,x}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that 
\begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=|\Target|}}\B_{x,\Target}(s)& \text{and}&&
\Prob_{x,\Target}(s) &= \frac{\B_{x,\Target}(s)}{\mathcal{Z}_{x,\Target}}.\label{def:distribution}\end{align}

It was established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the \GCContent distribution observed in sequences sampled with respect to the above distribution converges to a normal law of expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
Furthermore, it can easily be shown that the expected \GCContent is a continuous and strictly increasing monotonic  function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.

Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent. Such a value can be determined using any numerical recipe, and a simple bisection is used in our implementation. Furthermore, the concentration of the distribution, as asserted by the limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp. within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).

\subsection{Sampling the pseudo-Boltzmann ensemble}

Let us now describe a linear-time algorithm to sample sequences at random in the pseudo-Boltzmann distribution. This algorithm follows the general principles of the recursive approach to random generation~\cite{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\cite{Ding2003}. The algorithm starts by precomputing the partition function restricted to each valid interval, and then performs a series of stochastic backtracks, using precomputed values to decide on the probability of each alternative.

\subsubsection{Precomputation}
Let us describe a dynamic programming algorithm to compute $\Z{i,j}{a,b}$, the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.

Since the only sequence over an empty interval is the empty sequence, having energy $0$, one has
\begin{equation}
	\forall i \in [0,n-1]:\, \Z{i+1,i}{a,b}=1.
	\label{eq:Z_in}
\end{equation}

The recursion consists in four terms, depending on the base-pairing status and context of $(i,j)$:
\begin{equation*}
	\Z{i,j}{a,b}:=\left\{
  \begin{array}{ll}
  		\displaystyle
      \sum_{a'\in \B}  
      x^{\gc(a')}
      \cdot\Z{i+1,j}{a',b} &\begin{array}{@{}l@{}}\text{If }\Target_{i}=-1\\ \text{\relsize{-1}Position $i$ unpaired,}\end{array}\\
      \displaystyle
      \sum_{a',b'\in \B^2}
			 x^{\gc(a'.b')}
			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 \cdot \Z{i+1,j-1}{a',b'}&
\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1\\ \text{\relsize{-1}$(i,j)$ base-paired, stacked onto $(i$-$1,j$+$1)$,}\end{array}
			 \\
			 \displaystyle
      \sum_{a',b'\in \B^2}
      x^{\gc(a'.b')}
			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
      \cdot\Z{i+1,k-1}{a',b'}
      \cdot\Z{k+1,j}{b',b} & 
\begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j\\
\text{\relsize{-1} Junction or bulge $\Rightarrow$ split, no stacking.}
\end{array}
\\
      0 &\text{Otherwise}
	\end{array}\right.
\label{eq:Z_rec}
\end{equation*}
It is worth mentioning that, despite the appearances, this recurrence can be computed in linear $\Theta(|\Target|)$ time and space. Indeed, one can show that position $j$ is entirely determined by $i$ within subsequent recursive calls.

More precisely, let us consider the following predicate: {\em Let $(i',j')$ be the base-pair such that $i'<i<j'$ while minimizing $i-i'$, then TODO!!!!

$j$ is the position preceding the end of the closest base-pair nesting position $i$ if such a base-pair exists, or the last position in the sequence otherwise.} Obviously, the predicate holds for $[0,n-1]$. Moreover, if $i$ is unpaired, then $i+1$ and $i$

%The cases can be broken down as follows:
%\begin{description}
%\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
%any sequence consists in a nucleotide $a'$ at position  $i$, 
%followed by a sequence over $[i+1,j]$ and having flanking nucleotides $a'$ and $b$.
%\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$.
%Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
%then a stacking energy contribution is added. 
%\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
%only term contributing directly to the energy is the nucleotide distribution $\Prob(\cdot)$. 
%Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
%\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
%\end{description}

\subsubsection{Stochastic backtrack}
Once the partition functions have been computed, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, as described in Algorithm~\ref{alg:back}. 
\begin{algorithm}[t]
	\SetAlgoLined
\SetKwFunction{Backtrack}{Backtrack}
\SetKwFunction{Random}{Random}
	rand $\leftarrow$ \Random$\left(\Z{i,j}{a,b}\right)$\;
	k $\leftarrow S_i$\;
	\If(\tcp*[f]{Unpaired}){$k = -1$}{
		\For{$a'\in\B$}{
			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{i+1,j}{a',b}$\;
			\lIf{$\text{rand}<0$}\Return{$a'.\Backtrack\left((i+1,j),(a',b),S\right)$}\;
		}
	}
	\Else{
		\If(\tcp*[f]{Stacking pairs}){$k=j \land S_{i-1}=j+1$}{
			\For{$(a',b')\in\B\times\B$}{
				rand $\leftarrow$ rand $-
			 	x^{\gc(a'.b')}
			 	\cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
				\If{$\text{rand}<0$}{
					\Return{$a'.\Backtrack\left((i+1,j-1),(a',b'),S\right). b'$}\;		
				}
			}
		}
		\Else{
			\If(\tcp*[f]{Paired, not stacked}){$\Target_i=k \land i < k \leq j$}{
				\For{$(a',b')\in\B\times\B$}{
					rand $\leftarrow$ rand $-	
       x^{\gc(a'.b')}
					\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
  	    	\cdot\Z{i+1,k-1}{a',b'}
	    	  \cdot\Z{k+1,j}{b',b}$\;	
 					\If{$\text{rand}<0$}{
						\Return{$a'
						.\Backtrack\left((i+1,k-1),(a',b'),S\right)
						.b'
						.\Backtrack\left((k+1,j),(b',b),S\right)
						$}\;	
					}
				}
		 }
    \Else {\Return{$\varepsilon$};	\tcp*[f]{Empty sequence}}
		}
	}
\caption{\protect\Backtrack$\left((i,j),(a,b),S\right)$\label{alg:back}}
\end{algorithm}

\subsection{Overall complexity}
